### 2021-06-14 abstractESSLLI21/abstractESSLLI1.tex
abstract

\normalsize
\noindent {[}2{]} show that significant biases---thus measured---
similar to the ones discovered by IAT can be discovered in word
embeddings. {[}5{]} extends the methodology to a multilingual and
cross-lingual setting; a similar methodology is employed by {[}3{]}, who
use word embeddings trained on corpora from different decades. {[}6{]}
modify WEAT to a multi-class setting, introducing Mean Average Cosine
(MAC) similarity as a measure of bias (in fact, in the paper they report
distances rather than similarities). Let \(T = \{t_1, \dots, t_k\}\) be
a class of protected word embeddings, and let each \(A_j\in A\) be a set
of attributes stereotypically associated with a protected word). Then:

@@

We build analogous models for other topic groups (race, gender), and
different embeddings (we use three embeddings:
GoogleNews-vectors-negative300, and Reddit Corpus, and the hard-debiased
Reddit embedding used by {[}6{]}). The general conclusion is that from
this bayesian perpective the situation is much less obvious. The word
list sizes are small, sample sizes are small, and so posterior density
intervals are wide. Some stronger bias can be discerned in the gender
case. Moreover, sometimes the differences between associated, different
and human, are not very impressive. The cost of debiasing turns out
sometimes to be that neutral human predicates get closer to protected
classes (in religion), change in gender after debiasing is really minor
with zero still out out of HPDI range, and debiasing improvement for
race is a small improvement achieved at the price of moving neutral
terms closer to protected words (there is no space to include these
results in this abstract).

>>>>

\normalsize
\noindent {[}2{]} show that significant biases---thus measured---
similar to the ones discovered by IAT can be discovered in word
embeddings. {[}5{]} extend the methodology to a multilingual and
cross-lingual setting; a similar methodology is employed by {[}3{]}, who
use word embeddings trained on corpora from different decades. {[}6{]}
modify WEAT to a multi-class setting, introducing Mean Average Cosine
(MAC) similarity as a measure of bias (in fact, in the paper they report
distances rather than similarities). Let \(T = \{t_1, \dots, t_k\}\) be
a class of protected word embeddings, and let each \(A_j\in A\) be a set
of attributes stereotypically associated with a protected word). Then:

@@

We build analogous models for other topic groups (race, gender), and
different embeddings (we use three embeddings:
GoogleNews-vectors-negative300, and Reddit Corpus, and the hard-debiased
Reddit embedding used by {[}6{]}). The general conclusion is that from
this bayesian perpective the situation is much less obvious. The word
list sizes are small, sample sizes are small, and so posterior density
intervals are wide. Some stronger bias can be discerned in the gender
case. Moreover, sometimes the differences between associated, different
and human, are not very impressive. The cost of debiasing turns out
sometimes to be that neutral human predicates get closer to protected
classes (in religion), change in gender after debiasing is really minor
with zero still out of HPDI range, and debiasing improvement for race is
a small improvement achieved at the price of moving neutral terms closer
to protected words (there is no space to include these results in this
abstract).
